---
title: "Interactive Data Visualization in R"
author: "Lingge Li"
date: "3/20/2016"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Packages

ggplot 

- general data visualization package

ggmap

- built on top of ggplot for maps

shiny

- R based web framework

## Motivating example

http://www.calands.org/map

## Layered grammar of graphics

- ggplot2 follows a specific grammar of graphics

\includegraphics[width=300pt]{layer.png}

*Example taken from Hadley Wickham's book*

http://vita.had.co.nz/papers/layered-grammar.pdf

## How to make a plot

- Geometric objects (geom)
- Aesthetic mapping (aes)
- Statistical transformation (stat)
- Scales and coordinate system

## Geoms

- Wide range of geometric objects from points to complex shapes

- geom_point, geom_line, geom_polygon...

- Multiple geometric objects on the same plot with +


## Aesthetics

- Coordinate positions (always needed)

- Colour, fill, shape, size...

## Data + mapping

- aes() maps a dataframe to geom

- Each geom can have its own mapping

```{r echo=TRUE, eval=FALSE}
geom_point(data, aes(x, y))
```

## Points

```{r echo=FALSE, eval=TRUE}
library(ggplot2)
x <- c(1, 2, 3, 4, 2.5)
y <- c(3, 0, 1, 4, 4)
variable <- c(1, 2, 1, 2, 2)
example <- data.frame(x, y, variable)
```

```{r echo=TRUE, eval=TRUE, fig.width=4, fig.height=2, fig.align='center'}
ggplot(data=example) +
  geom_point(aes(x=x, y=y, colour=as.factor(variable)))
```

## Shape

```{r echo=TRUE, eval=TRUE, fig.width=4, fig.height=2, fig.align='center'}
ggplot(data=example) +
  geom_point(aes(x=x, y=y, colour=as.factor(variable))) +
  geom_polygon(aes(x=x, y=y), colour='black', fill=NA)
```

## Polygons

```{r echo=FALSE, eval=TRUE}
x <- c(5, 6, 4)
y <- c(4, 3, 1)
variable <- c(3, 3, 3)
triangle <- data.frame(x, y, variable)
```

```{r echo=TRUE, eval=TRUE, fig.width=4, fig.height=2, fig.align='center'}
example$group <- 1
triangle$group <- 2
both <- rbind(example, triangle)
ggplot(data=both) +
  geom_point(aes(x=x, y=y, colour=as.factor(variable))) +
  geom_polygon(aes(x=x, y=y, group=group), colour='black', fill=NA)
```

## States dataframe

```{r echo=TRUE, eval=TRUE, message=FALSE}
library(ggmap)
library(maps)

states <- map_data('state')
head(states)
```

## USA

```{r echo=TRUE, eval=TRUE, message=FALSE, fig.width=4, fig.height=3, fig.align='center'}
ggplot(data=states) + 
  geom_polygon(aes(x=long, y=lat, group=group, fill=region), show.legend=FALSE)
```

## California

```{r echo=TRUE, eval=TRUE, message=FALSE, fig.width=3, fig.height=3, fig.align='center'}
california <- states[states$region == 'california', ]
ggplot(data=california) + 
  geom_polygon(aes(x=long, y=lat), fill='orange', color='red')
```

## Custom scale

```{r echo=TRUE, eval=TRUE, message=FALSE, fig.width=4, fig.height=3, fig.align='center'}
ggplot(data=california) + 
  geom_polygon(aes(x=long, y=lat), fill='orange', color='red') +
  scale_x_continuous(limits=c(-125, -110)) +
  scale_y_continuous(breaks=c(34, 37), labels=c('dude', 'hella'))
```

## Details

- Theme 

- http://docs.ggplot2.org/current/theme.html

## ggmap

https://github.com/dkahle/ggmap

## get_map

```{r echo=TRUE, eval=TRUE, message=FALSE, fig.width=4, fig.height=3, fig.align='center'}
cali <- get_googlemap(center=c(lon=-120, lat=37.5), maptype='satellite', zoom=6)
ggmap(cali)
```

## Shapefile

```{r echo=TRUE, eval=FALSE}
setwd('/Users/linggeli/Downloads/CAPD')
CAPD <- readOGR('.', 'CPAD_2016b1_SuperUnits')
CAPD <- spTransform(CAPD, CRS("+proj=longlat +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))
proj4string(CAPD) <- CRS("+proj=longlat +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")
state <- CAPD[CAPD$MNG_AG_LEV == 'State', ]
state <- fortify(state)
state.low <- unique(data.frame(long=round(state$long, 2), lat=round(state$lat, 2), group=state$group))
```

## Protected areas

```{r echo=FALSE, eval=TRUE, message=FALSE, fig.width=4, fig.height=3, fig.align='center'}
load('/Users/linggeli/Downloads/CAPD/protected.Rda')
ggmap(cali) + 
  geom_polygon(data=protected, aes(x=long, y=lat, group=group, fill=level), colour='NA', alpha=0.8) +
  scale_fill_manual(values=c('Federal'='green', 'State'='orange')) +
  labs(title='Protected Lands in California') + 
  theme(legend.position=c(0.2, 0.2)) +
  theme(plot.title=element_text(hjust=0.5))
```

## What is Shiny

- A framework for building web applications

- Best for interactive data visualization

- Apps for exploratory analysis

http://shiny.datascience.uci.edu/UCIDataScienceInitiative/ClimateActionShiny/

http://shiny.datascience.uci.edu/uciMetropolitanFutures/employment_centers/

## Server and UI

- ui.R has everything you see

input widgets, plots, tables...

- server.R does the work

```{r echo=TRUE, eval=FALSE}
shinyServer(function(input, output) {
  
})

shinyUI(fluidPage(
  
))
```

## Input

- Input handled by specific widgets

- Each input has an id

- Access input value with input$id

http://shiny.rstudio.com/gallery/widget-gallery.html

## Output

- Output rendered in server with output$id

- Then displayed in ui

```{r echo=TRUE, eval=FALSE}
library(shiny)
shinyServer(function(input, output) {
  output$histogram <- renderPlot({
    hist(faithful$eruptions)
  })
})

shinyUI(fluidPage(
  plotOutput(outputId='histogram')
))
```

## Several types of output

- plotOutput (imageOutput)
- tableOutput (dataTableOutput)
- textOutput (html)
- verbatimTextOutput (console)
- htmlOutput (uiOutput)

## Reactive environment

- Triggered when input changes

- Output changes accordingly

- Rendering functions reactive

## Example

```{r echo=TRUE, eval=FALSE}
library(shiny)
library(ggplot2)
library(ggmap)

shinyServer(function(input, output) {
  
  load('/Users/linggeli/Downloads/CAPD/parks.Rda')
   
  output$distPlot <- renderPlot({
    cali <- get_googlemap(center=c(lon=-120, lat=37.5), maptype='hybrid', zoom=6)
    ggmap(cali) +
      geom_polygon(data=parks, aes(x=long, y=lat, group=group), colour='NA', fill='green', alpha=0.8)
  })
  
})
```

```{r echo=TRUE, eval=FALSE}
shinyUI(fluidPage(
  plotOutput('distPlot', height='640px')
))
```

## Slide to zoom

```{r echo=TRUE, eval=FALSE}
shinyUI(fluidPage(
  sliderInput('res', label='Zoom', min=6, max=9, value=6),
  plotOutput('distPlot', height='640px')
))
```

```{r echo=TRUE, eval=FALSE}
output$distPlot <- renderPlot({
    cali <- get_googlemap(center=c(lon=-120, lat=37.5), maptype='hybrid', zoom=input$res)
    ggmap(cali) +
      geom_polygon(data=parks, aes(x=long, y=lat, group=group), colour='NA', fill='green', alpha=0.8)
  })
```

## Double click to move

```{r echo=TRUE, eval=FALSE}
location <- reactiveValues(x=-120, y=37.5)
  
  observeEvent(input$plot_dblclick, {
    current <- input$plot_dblclick
    location$x <- current$x
    location$y <- current$y
  })
```

## Layout

- Fluid grid layout (similar to bootstrap)

- 12 columns every row

- Tabset

\includegraphics[width=300pt]{grid.png}

## Other packages

- Widgets for Javascript data visualization

http://www.htmlwidgets.org/

## Resources

- Documentation 

http://docs.ggplot2.org/current/

- Gallery with source code

http://shiny.rstudio.com/gallery/

- Cheatsheets

https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf
http://shiny.rstudio.com/images/shiny-cheatsheet.pdf